Review Paper Draft

### 2020-07-07 main.tex
4 and 5 edits

%------------------------------%
\subsection{LMM-Lasso}
LMM-Lasso models $\bZ\bgamma$ as a random effect assuming 
$$\bgamma \sim N\left(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p \right).$$  Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.This results in the LMM regression model
\begin{equation}
    \label{eqn:lmm_reg}
    \by = \Int + \bX\bbeta + \bX\bgamma+ \beps,
\end{equation}
where we assume $\beps \independent \bX\bgamma$. Rather than assuming anything about $\bgamma$ directly, an equivalent representation lets $\bu=\bX\bgamma$ such that $\bu \sim N(\boldsymbol{0}, \sigmagg \bK)$, and yields the equivalent model
\begin{equation}
    \label{eqn:lmm_reg_u}
    \by = \Int + \bX\bbeta + \bu + \beps.
\end{equation}
Regardless of representation, the LMM assumes $\by \sim N(\Int + \bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$. LMM-Lasso performs a weighted rotation of the original data, left-multiplying by $\bV^{-1/2}$ in a manner akin to generalized least squares to diagonalize the error term such that
\begin{equation}
\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \left(\Int +\bX \bbeta\right), \bI).
\end{equation}
\anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?}
\begin{comment}
For ease of notation, we incorporate the intercept into the design matrix and let $\bX' = \begin{bmatrix} \textbf{1} & \bX \end{bmatrix}$ and $\bbeta' = \begin{pmatrix} \mu & \bbeta \end{pmatrix}^T$ so that we can express this more compactly as 
\begin{equation}
\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \bX' \bbeta', \bI).
\end{equation}
\end{comment}
Using the spectral decomposition of $\bK$ described in equation \eqref{eqn:k}, we can gain further insight into the mechanism of this transformation. Note that $\bV^{-1}$ can be expressed using the following matrix factorization 
\begin{align*}
    \bV^{-1} &= (\sigmagg \bK + \sigmaee \bI)^{-1}\\
    &=(\sigmagg \bU \bD^2 \bU^T + \sigmaee \bU \bU^T)^{-1}\\
    &= (\bU (\sigmagg \bD^2 + \sigmaee \bI) \bU^T)^{-1}\\
    &=\bU (\sigmagg \bD^2 + \sigmaee \bI)^{-1} \bU^T.
\end{align*}
By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$ and $\byT = \bU^T \left(\by - \Int \right)$, where we center the transformed $\by$ to simplify notation and computation.  $\bW = \text{diag}\{w_1, ... w_n\}$ where $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \boldsymbol{D}^2 + \sigmaee \bI)^{-1}$. Together, this yields the objective function
\begin{equation}
\label{eqn:lmm_obj}
Q(\bbeta | \bXT, \byT, \bW) = || \bW (\byT - \bXT \bbeta)||_2^2 + \lambda || \bbeta ||_1.
\end{equation}
An advantage to modeling unobserved confounding using a random effect, as opposed to fixed effects as in PC-Lasso, is that this requires the estimation of substantially fewer parameters: $p + 2$ for LMM-Lasso-Rakitsch, and $p + 3$ for LMM-Lasso-ggmix.

@@

We have so far used the term $\bW$ to describe the general LMM-Lasso method. However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix. LMM-Lasso-Rakitsch \citep{Rakitsch2012} uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$. These variance components and the corresponding weights are estimated once under the assumption of a null model. A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \citep{glmnet} or \texttt{ncvreg} \citep{ncvreg}. LMM-Lasso-ggmix \citep{bhatnagar2019simultaneous} proposes iteratively updating $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm. This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \citep{ggmix, bhatnagar2019simultaneous}. We postpone further discussion of the differences between these methods to the results section.

@@

%------------------------------%
\subsection{LMM-Lasso outperforms PC-Lasso when simulated subpopulations are small}
%------------------------------%

Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compared these methods in the presence of fine and coarse subpopulation structures.

Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying environmental confounding, $\xi \in \{0, 0.2, 0.5,0.8\}$. Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the independent subpopulation simulated $\bX$ with environmental confounding simulated using a dichotomous-heterogeneous $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\gamma$ structures.

with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures. The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy. In the presence of a small number of large subpopulations, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding box straddling the 0 line. However, the majority of the box falls below 0, indicating the superior performance of PC-Lasso in many of the simulations. In the presence of many small subpopulations, however, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy. This is consistent with \citep{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models. As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations. 
\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with independent subpopulation data, $\eta = 0.5$, and dichotomous-discordant environmental effects.}
    \label{fig:big_vs_small}
\end{figure}

%------------------------------%
\subsection{LMM-Lasso outperforms PC-Lasso when $\gamma$ is heterogeneous}
%------------------------------%

We evaluate PC-Lasso and LMM-Lasso in the presence of large and small subpopulations as well as distinct coefficient structures. Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) for all simulations, broken down by both subpopulation and coefficient structure. The right-hand panels of \ref{fig:big_small_gamma} are consistent with figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of small subpopulations, regardless of coefficient linearity or homogeneity. The panel of figure \ref{fig:big_small_gamma} corresponding to large subpopulations with heterogeneous coefficients also appears to be a setting where LMM-Lasso consistently outperforms PC-Lasso, regardless of coefficient linearity. Only in the setting of large supopulations with homogeneous coefficients does PC-Lasso outperform LMM-Lasso. In line with our initial intuition, PC-Lasso performs best within this setting when coefficients are linear in nature.

@@

Section~\ref{sec:rak-ggmix} briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix. As we will see, the two generally perform quite similarly. Thus, the preceding results, although based on the LMM-Lasso-Rakitsch implementation, are representative of either LMM-Lasso method.

The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding. However, given the generally strong performance of LMM-Lasso, we also wish to briefly discuss the key differences between these two ways of implementing it, and provide guidance for potential users.

LMM-Lasso-Rakitsch estimates the variance components once, under the assumption of a null model. LMM-Lasso-ggmix seeks to improve upon this by re-estimating the variance components and $\bbeta$ iteratively. Although this seems a reasonable approach, we have been unable to find a scenario where LMM-Lasso-ggmix outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$ or $\eta$.

@@

In addition, we have found that LMM-Lasso-ggmix performs poorly in a low-dimensional setting due to the ill-conditioning of the RRM. Given its intended use for the analysis of high-dimensional data, this fact could be of little consequence to practitioners. However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in figure \ref{fig:eta_beta_mse}, this is also a setting in which the computational time required may be prohibitive for practical applications. Figure \ref{fig:time} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes. The horizontal dashed line at 1 indicates both methods require the same computing time. Despite its comparable performance in accurately estimating $\hat{\bbeta}$, LMM-Lasso-ggmix takes approximately 71 to 9 times as long as LMM-Lasso-Rakitsch. Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion} \label{sec:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this work, we have reviewed the concept of population stratification and defined explicitly the environmental (i.e., non-genetic) mechanism by which it may confound the genotype-phenotype relationship. We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences. In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data: PC-Lasso and LMM-Lasso.

We have detailed the assumptions made by PC-Lasso and LMM-Lasso, which model confounding influences as fixed and random effects, respectively. We have also detailed the ways in which simulation schemes to evaluate these methods may fail to accurately capture the effects of environmental heterogeneity. In turn, we propose and utilize a simulation approach in which the estimation bias introduced by environmental heterogeneity may be evaluated. We then compare the estimation accuracy of a naive lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.

Unsurprisingly, we find that PC-Lasso and LMM-Lasso outperform the naive lasso whenever environmental confounding effects are present. However, in the absence of such effects, PC-Lasso may hinder estimation accuracy and perform more poorly than even the naive lasso. This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability. Additionally, we find that LMM-Lasso outperforms PC-Lasso when (1) small genetic subpopulations are present which correspond to environmental heterogeneity and (2) when confounding effects are nonlinear and heterogeneous in nature. In the presence of large subpopulations which correspond to homogeneous, linear environmental effects, PC-Lasso and LMM-Lasso perform comparably in estimation accuracy. While this scenario may be plausible, given that LMM-Lasso is more robust to subpopulation sizes and nonlinear effects, it may present a 'safer choice' for practitioners seeking to obtain unbiased estimates when the exact nature of population structure and corresponding environmental heterogeneity are unknown.

We considered two implementations of the LMM-Lasso method: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, which differ in their variance estimation procedures and computational implementations. Although we found that both methods perform similarly in terms of estimation accuracy in many scenarios, LMM-Lasso-ggmix has several distinct disadvantages. Firstly, it suffers from instability in low dimensions. Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably in high dimensions, LMM-Lasso-ggmix takes approximately \anna{x} times longer due to its iterative estimation procedure. Thirdly, while the additional time required of LMM-Lasso-ggmix could be worthwhile if it resulted in a more accurate estimation of $\eta$, or narrow-sense heritability, we have been unable to identify a scenario where its narrow-sense heritability estimates are superior to those of LMM-Lasso-Rakitsch. We have implemented the LMM-Lasso-Rakitsch method using the equivalent but more interpretable $\eta$ paramaterization in the \texttt{R} package, \texttt{penalizedLMM (final name tbd)}, available on the author's github page: \url{https://github.com/areisett/penalizedLMM}.

@@

Even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods. Because of this, bias-reduction techniques for penalized LMMs may be an attractive avenue for future research.

>>>>

%------------------------------%
\subsection{LMM-Lasso}
LMM-Lasso models $\bZ\bgamma$ as a random effect assuming 
$$\bgamma \sim N\left(\boldsymbol{0}, \frac{\sigmagg}{p} \bI_p \right).$$  Typically, $\bZ$ is not observed directly, and a common convention is to assume $\bZ = \bX$ \citep{wang2018multiplex, lippert2011fast, yang2014advantages}.This results in the LMM regression model
\begin{equation}
    \label{eqn:lmm_reg}
    \by = \Int + \bX\bbeta + \bX\bgamma+ \beps,
\end{equation}
where we assume $\beps \independent \bX\bgamma$. Rather than assuming anything about $\bgamma$ directly, an equivalent representation lets $\bu=\bX\bgamma$ such that $\bu \sim N(\boldsymbol{0}, \sigmagg \bK)$, and yields the equivalent model
\begin{equation}
    \label{eqn:lmm_reg_u}
    \by = \Int + \bX\bbeta + \bu + \beps.
\end{equation}
Regardless of representation, the LMM assumes $\by \sim N(\Int + \bX \bbeta, \bV)$, where $\bV = \sigmagg \textbf{K} + \sigmaee \textbf{I}$. LMM-Lasso performs a weighted rotation of the original data, left-multiplying by $\bV^{-1/2}$ in a manner akin to generalized least squares to diagonalize the error term such that
\begin{equation}
\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \left(\Int +\bX \bbeta\right), \bI).
\end{equation}
\anna{...however, as Karl Rohe paper shows, GLS does not completely translate to penalized regression...comment on this?}
\begin{comment}
For ease of notation, we incorporate the intercept into the design matrix and let $\bX' = \begin{bmatrix} \textbf{1} & \bX \end{bmatrix}$ and $\bbeta' = \begin{pmatrix} \mu & \bbeta \end{pmatrix}^T$ so that we can express this more compactly as 
\begin{equation}
\bV^{-\frac{1}{2}}\by \sim N(\bV^{-\frac{1}{2}} \bX' \bbeta', \bI).
\end{equation}
\end{comment}
Using the spectral decomposition of $\bK$ described in equation \eqref{eqn:k}, we can gain further insight into the mechanism of this transformation. Note that $\bV^{-1}$ can be expressed using the following matrix factorization 
\begin{align*}
    \bV^{-1} &= (\sigmagg \bK + \sigmaee \bI)^{-1}\\
    &=(\sigmagg \bU \bD^2 \bU^T + \sigmaee \bU \bU^T)^{-1}\\
    &= (\bU (\sigmagg \bD^2 + \sigmaee \bI) \bU^T)^{-1}\\
    &=\bU (\sigmagg \bD^2 + \sigmaee \bI)^{-1} \bU^T.
\end{align*}
By incorporating these decompositions, the transformed data can be expressed as $\bW \bXT$ and $\bW \byT$, respectively, where $\bXT = \bU^T \bX$ and $\byT = \bU^T \left(\by - \Int \right)$, where we center the transformed $\by$ to simplify notation and computation.  $\bW = \text{diag}\{w_1, ... w_n\}$ where $w_i$ corresponds to the square root of the $i$th element of the diagonal matrix $(\sigmagg \bD^2 + \sigmaee \bI)^{-1}$. Together, this yields the objective function
\begin{equation}
\label{eqn:lmm_obj}
Q(\bbeta | \bXT, \byT, \bW) = || \bW (\byT - \bXT \bbeta)||_2^2 + \lambda || \bbeta ||_1.
\end{equation}
An advantage to modeling unobserved confounding using a random effect, as opposed to fixed effects as in PC-Lasso, is that this requires the estimation of only $p + 3$ parameters.

@@

We have so far used the term $\bW$ to describe the general LMM-Lasso method. However, the form of $\bW$ and method of variance component estimation differs between LMM-Lasso-Rakitsch and LMM-Lasso-ggmix. LMM-Lasso-Rakitsch \citep{Rakitsch2012} uses a two-step procedure for estimating the variance components, $\sigmaee$ and $\delta := \sigmaee / \sigmagg$. These variance components and the corresponding weights are estimated once under the assumption of a null model. A Lasso model can then be fit on the transformed data to estimate $\boldsymbol{\beta}$ using standard software for coordinate descent, such as \texttt{glmnet} \citep{glmnet} or \texttt{ncvreg} \citep{ncvreg}. LMM-Lasso-ggmix \citep{bhatnagar2019simultaneous} proposes iteratively updating $\widehat{\boldsymbol{\beta}}$ and the variance components via a block coordinate descent algorithm. This iterative update scheme necessitates the use of the corresponding \texttt{R} package \texttt{ggmix} \citep{ggmix, bhatnagar2019simultaneous}. We postpone further discussion of the differences between these methods to section \ref{sec:results}.

@@

%------------------------------%
\subsection{LMM-Lasso outperforms PC-Lasso in the presence of fine subpopulation structure}
%------------------------------%

Given the superior performance of PC-Lasso and LMM-Lasso compared to the naive lasso in the presence of environmental confounding, we next compare these methods in the presence of fine and coarse subpopulation structures. Phenotypes were simulated using a 1:1 SNR, $\eta = 0.5$, and varying levels of environmental confounding, $\xi \in \{0, 0.2, 0.5,0.8\}$. Figure \ref{fig:big_vs_small} summarizes the results of these simulations for the independent subpopulation simulated $\bX$ with environmental confounding generated using a dichotomous-discordant $\bgamma$ structure, though qualitatively similar results were observed for the other data types and $\gamma$ structures. The y-axis displays the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures. The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy.

In the presence of coarse population structure, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding boxes straddling the 0 line. However, the majority of these boxes fall above 0, particularly for lower $\eta$ values, indicating the superior performance of LMM-Lasso. In the presence of fine population structure, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy. This is consistent with \citep{hoffman2013correcting}, who showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models. As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform reasonably well in the coarse subpopulation setting, but fall short in the presence of fine structure. 
\begin{figure}[H]
    \centering
    % \includegraphics[scale = 0.9]{figures/mse_diff_subpops}
     \includegraphics[scale = 0.9]{figures/mse_diff_subpops.png}
    \caption{Relative performance of LMM-Lasso, PC-Lasso for fine and coarse subpopulation structures and varying levels of environmental confounding with independent subpopulation data, $\eta = 0.5$, and dichotomous-discordant environmental effects.}
    \label{fig:big_vs_small}
\end{figure}

%------------------------------%
\subsection{LMM-Lasso outperforms PC-Lasso when environmental effects are discordant}
%------------------------------%

In addition to evaluating PC-Lasso and LMM-Lasso in the presence of coarse and fine subpopulation structure, we also consider distinct patterns of environmental effects. Figure \ref{fig:big_small_gamma} shows the difference in MSE (PC-Lasso - LMM-Lasso) broken down by both subpopulation and coefficient structure. Results are shown for phenotypes generated with $\eta = 0.5$, $\xi = 0.8$, and the independent subpopulation simulated $\bX$. The right-hand panels of figure \ref{fig:big_small_gamma} are consistent with figure \ref{fig:big_vs_small}, showing superior LMM-Lasso performance in the presence of fine subpopulation structure, regardless of coefficient magnitude or concordance. The upper left hand panel of figure \ref{fig:big_small_gamma} corresponds to coarse subpopulation structure with discordant coefficients and also appears to be a setting where LMM-Lasso outperforms PC-Lasso, regardless of coefficient magnitude, although there is some overlap with PC-Lasso. The lower left hand panel of figure \ref{fig:big_small_gamma} shows results for the setting of coarse supopulation structure with concordant coefficients. In this setting, PC-Lasso and LMM-Lasso perform comparably in the presence of exponential-concordant environmental effects, while PC-Lasso outperforms LMM-Lasso in the majority of dichotomous-concordant environmental effects. These findings are in line with our initial hypothesis that PC-Lasso will perform well in the presence of environmental effects that can be well-approximated by a linear combination of the observed data. When that is not the case, either due to the presence of fine subpopulation structure, nonlinear environmental effects, or genetic similarity that does not reflect similar environmental effects, LMM-Lasso exhibits estimation accuracy superior to that of PC-Lasso.

@@

In section \ref{sec:rak-ggmix} we briefly mentioned the existence of two LMM-Lasso variants: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix. As we will see, the two generally perform quite similarly. Thus, the preceding results, although based on the LMM-Lasso-Rakitsch implementation, are representative of both LMM-Lasso methods. The primary purpose of this review is to compare LMM-Lasso to other penalized regression methods with respect to correcting for population structure and environmental confounding. However, given the generally strong performance of LMM-Lasso, we also wish to discuss the key differences between these two implementations and provide guidance for potential users.

LMM-Lasso-Rakitsch estimates the variance components once, under the assumption of a null model. LMM-Lasso-ggmix seeks to improve upon this by re-estimating the variance components and $\bbeta$ iteratively. Although this seems a reasonable approach, we have been unable to find a scenario where LMM-Lasso-ggmix outperforms the LMM-Lasso-Rakitsch in estimation accuracy of $\bbeta$. Figure \ref{fig:eta_beta_mse} shows the MSE for estimation of both $\bbeta$ and $\eta$ in the presence of varying $\eta$ values and sample sizes, $n$, for both LMM-Lasso-Rakitsch and LMM-Lasso-ggmix. The bottom panel of this figure shows a modest improvement in the estimation of $\eta$ for the LMM-Lasso-ggmix method, particularly in the presence of lower $\eta$ values. However, these apparent improvements in estimating $\eta$ do not improve the $\bbeta$ estimation accuracy of LMM-Lasso-ggmix. This is evidenced by the top panel of figure \ref{fig:eta_beta_mse}, which shows nearly identical performances of both methods in terms of $\bbeta$ MSE.

@@

In addition, we have found that LMM-Lasso-ggmix performs poorly in a low-dimensional setting due to the ill-conditioning of the RRM. Given its intended use for the analysis of high-dimensional data, this fact could be of little consequence to practitioners. However, although its estimation accuracy in high dimensions is on par with that of LMM-Lasso-Rakitsch, as shown in figure \ref{fig:eta_beta_mse}, this is also a setting in which the computational time required may be prohibitive for practical applications. Figure \ref{fig:time} shows the ratio of computing time required for LMM-Lasso-ggmix compared to that of LMM-Lasso-Rakitsch, for various sample sizes. The horizontal dashed line at 1 indicates both methods require the same computing time. Despite its comparable performance in accurately estimating $\hat{\bbeta}$, LMM-Lasso-ggmix takes requires as much as 75 times the computing time of LMM-Lasso-Rakitsch. Thus, LMM-Lasso-Rakitsch seems to be a more attractive and scalable method than LMM-Lasso-ggmix.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion} \label{sec:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%------------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this work, we have reviewed the concept of population stratification and explicitly defined the environmental (i.e., non-genetic) mechanism by which it may confound the genotype-phenotype relationship. We have provided a broad outline of existing methods for ameliorating the effects of these unobserved confounding influences. In addition, we have provided a thorough review of two popular approaches for correcting for population structure in genetic data: PC-Lasso and LMM-Lasso.

We have detailed the assumptions made by PC-Lasso and LMM-Lasso, which model confounding influences as fixed and random effects, respectively. We have also described the ways in which simulation schemes to evaluate these methods may fail to accurately capture the effects of environmental heterogeneity. In turn, we propose and utilize a simulation approach in which the estimation bias introduced by environmental heterogeneity may be evaluated. We then compare the estimation accuracy of a naive lasso, PC-Lasso, and LMM-Lasso in various confounding scenarios.

Unsurprisingly, we find that PC-Lasso and LMM-Lasso outperform the naive lasso whenever environmental confounding effects are present. However, in the absence of such effects, PC-Lasso may hinder estimation accuracy and perform more poorly than even the naive lasso. This is due to the fact that the principal components are unpenalized and thus included in the final model, even when they do more harm than good in explaining phenotypic variability. Additionally, we find that LMM-Lasso outperforms PC-Lasso when (1) fine subpopulation structure is present and (2) when confounding effects are nonlinear and discordant in nature. In the presence of coarse subpopulation structure which corresponds to concordant, dichotomous environmental effects, PC-Lasso may outperform LMM-Lasso in estimation accuracy. However, LMM-Lasso is more robust to subpopulation structure as well as nonlinear and discordant environmental effects. Therefore, LMM-Lasso is a more robust method for practitioners seeking to obtain unbiased estimates when the exact nature of the population structure and corresponding environmental heterogeneity are unknown.

Finally, we considered two implementations of the LMM-Lasso method: LMM-Lasso-Rakitsch and LMM-Lasso-ggmix, which differ in their variance estimation procedures and computational implementations. Although we found that both methods perform similarly in terms of $\bbeta$ estimation accuracy, LMM-Lasso-ggmix has several drawbacks. Firstly, it suffers from instability in low dimensions. Secondly, although LMM-Lasso-Rakitsch and LMM-Lasso-ggmix perform comparably in high dimensions, LMM-Lasso-ggmix requires as much as \anna{75 - check} times the computing time of LMM-Lasso-Rakitsch due to its iterative estimation procedure. Thirdly, while the additional time required of LMM-Lasso-ggmix may be worthwhile if it yielded more accurate estimates of $\eta$, or narrow-sense heritability, we have been unable to identify a scenario where its narrow-sense heritability estimates are greatly superior to those of LMM-Lasso-Rakitsch. We have implemented the LMM-Lasso-Rakitsch method using the equivalent, but more interpretable, $\eta$ paramaterization in the \texttt{R} package, \texttt{penalizedLMM (final name tbd)}, available on the author's github page: \url{https://github.com/areisett/penalizedLMM}.

@@

It is worth nothing that even in scenarios where LMM-Lasso outperforms other methods, its estimation accuracy still suffers due to the bias inherent in penalized regression methods. Because of this, bias-reduction techniques for penalized LMMs may be an attractive avenue for future research.
